---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-"
)
```

# ungeviz

Tools for visualizing uncertainty with ggplot2, written by Claus O. Wilke

This package is in the very early stages of development. Most things are broken. However, the bootstrapping functions may already be useful for generating hypothetical outcomes plots.

The package name comes from the German word "Ungewissheit", which means uncertainty.

## Installation

```{r eval = FALSE}
devtools::install_github("clauswilke/ungeviz")
```


## Bootstrapping {#bootstrap}

The functions `bootstrap_summarize()` and `bootstrap_do()` are drop-in equivalents for dplyr's `summarize()` and `do()` but perform the respective action multiple times on bootstrapped data. Similarly, the function `bootstrap_collect()` generates a bootstrapped data frame and the function `bootstrapper()` generates a bootstrapping object that can be used instead of data in ggplot2 layers.

```{r diamonds-mean-anim, message = FALSE, warning = FALSE}
library(tidyverse)
library(ungeviz)
library(gganimate)

diamonds %>% group_by(cut, color, clarity) %>%
  bootstrap_summarize(20, mean_price = mean(price)) %>%
  ggplot(aes(color, mean_price, color = clarity)) +
    geom_point() + facet_wrap(~cut) +
    theme_bw() + 
    # `.bootstrap` is a generated column indicating the bootstrap
    # draw a data row belongs to 
    transition_states(.bootstrap, 1, 1)
```

```{r mtcars-smooth-anim}
mtcars %>% bootstrap_collect(20) %>%
  ggplot(aes(mpg, hp)) + 
    geom_smooth(
      data = mtcars,
      method = "gam", formula = y ~ s(x, k = 3),
      color = NA
    ) + 
    geom_point(data = mtcars) +
    geom_smooth(
      method = "gam",
      formula = y ~ s(x, k = 3),
      se = FALSE
    ) + 
    theme_bw() +
    transition_states(.bootstrap, 1, 1)
```

```{r bluejays-lm-anim}
data(BlueJays, package = "Stat2Data")

# set up bootstrapping object that generates 20 bootstraps
# and groups by variable `KnownSex`
bs <- ungeviz::bootstrapper(20, KnownSex)

ggplot(BlueJays, aes(BillLength, Head, color = KnownSex)) +
  geom_smooth(method = "lm", color = NA) +
  geom_point(alpha = 0.3) +
  # `.row` is a generated column providing a unique row number
  # to all rows in the bootstrapped data frame 
  geom_point(data = bs, aes(group = .row)) +
  geom_smooth(data = bs, method = "lm", fullrange = TRUE, se = FALSE) +
  facet_wrap(~KnownSex, scales = "free_x") +
  scale_color_manual(values = c(F = "#D55E00", M = "#0072B2"), guide = "none") +
  theme_bw() +
  transition_states(.bootstrap, 1, 1) + 
  enter_fade() + 
  exit_fade()
```

## Visualizing uncertainty from fitted models {#fitted-models}

Some very early code exists to help visualizing uncertainty from fitted models, for example as confidence strips.

```{r}
library(broom)

df_model <- lm(mpg ~ disp + hp + qsec, data = mtcars) %>%
  tidy() %>%
  filter(term != "(Intercept)")

ggplot(df_model, aes(x = estimate, moe = std.error, y = term)) +
  stat_confidence_density(fill = "lightblue", height = 0.8, confidence = 0.68) +
  geom_point(aes(x = estimate), size = 3) +
  geom_errorbarh(aes(xmin = estimate - std.error, xmax = estimate + std.error), height = 0.5) +
  scale_alpha_identity() +
  xlim(-2, 1)
```

