---
title: "Supporting geoms and stats"
author: "Claus O. Wilke"
date: "`r Sys.Date()`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  fig.width = 4,
  comment = "#>"
)
```

The package provides a few supporting geoms and stats that can be generally helpful when generating uncertainty visualizations.

## Horizontal and vertical plines

When showing an individual value, such as a mean, we may want to show it as a vertical or horizontal line instead of a point. For this purpose, ungeviz provides horizontal and vertical point lines (plines), through the two geoms `geom_hpline()` and `geom_vpline()`. These geoms behave like points but can be styled like lines. They have the useful advantage over `geom_segment()` that they work with both continuous and discrete position scales.

```{r iris-hpline, message = FALSE}
library(ggplot2)
library(ungeviz)

ggplot(iris, aes(Species, Sepal.Length)) +
  geom_point(position = position_jitter(width = 0.3, height = 0), size = 0.5) +
  geom_hpline(aes(colour = Species), stat = "summary", width = 0.6, size = 1.5)
```

Plines have a length and a thickness. The length is specified as `width` and `height`, respectively, for the horizontal and vertical variants. The thickness is always specified as `size`. Note that `width` and `height` specify the total length of the pline, whereas the equally named parameters in `position_jitter()` specify only the linear extent in one direction. Therefore, in the above example, we use `width = 0.3` for `position_jitter()` and `width = 0.6` for `geom_hpline()`.

Plines can be combined with error bars to show parameter estimates and their uncertainties in a regression model.

```{r cacao-vpline}
library(dplyr)
library(forcats)
library(broom)
library(emmeans)

cacao_lumped <- cacao %>%
  mutate(
    location = fct_lump(location, n = 10)
  )
  
cacao_means <- lm(rating ~ location, data = cacao_lumped) %>%
  emmeans("location") %>%
  tidy() %>%
  mutate(location = fct_reorder(location, estimate))

ggplot(cacao_means, aes(x = estimate, y = location)) +
  geom_errorbarh(aes(xmin = estimate - std.error, xmax = estimate + std.error), height = 0.3) +
  geom_vpline(aes(x = estimate), size = 2, height = 0.7, color = "#D55E00") +
  xlim(2.8, 3.6) + 
  theme_minimal()
```

## Confidencde strips and confidence densities

Sometimes we may want to visualize the uncertainty distributions as colored bands that fade out (called confidence strips). Confidence strips can be generated with `stat_confidence_density()`. This stat takes as input a mean value (mapped to `x`), a margin of error (mapped to `moe`), and a confidence level corresponding to the margin of error (mapped to `confidence`). It uses these inputs to calculate the corresponding density function for a normal distribution, which it provides as new variable `stat(density)`. A `stat(ndensity)` is also provided, which is scaled so its maximum value is one. `stat_confidence_density()` automatically maps the normalized density to the `alpha` aesthetic, so that we can draw confidence strips simply by specifying a fill color.

```{r cacao-confidence-density}
ggplot(cacao_means, aes(x = estimate, y = location)) +
  stat_confidence_density(aes(moe = std.error), confidence = 0.68, fill = "#81A7D6", height = 0.7) +
  geom_errorbarh(aes(xmin = estimate - std.error, xmax = estimate + std.error), height = 0.3) +
  geom_vpline(aes(x = estimate), size = 2, height = 0.7, color = "#D55E00") +
  xlim(2.8, 3.6) + 
  theme_minimal()
```

There is one important caveat for `stat_confidence_density()`: It will generally not work for non-linear or transformed scales, because it uses the margin of error to fit a normal distribution and the moe is not modified by these scale transformation. I would recommend against using `stat_confidence_density()` with anything but a standard linear scale.


```{r cacao-confidence-ridgeline}
library(ggridges)

ggplot(cacao_means, aes(x = estimate, y = location)) +
  stat_confidence_density(
    aes(moe = std.error, height = stat(density)), geom = "ridgeline",
    confidence = 0.68, fill = "#81A7D6", alpha = 0.8, scale = 0.08,
    min_height = 0.1
  ) +
  geom_vpline(aes(x = estimate), size = 2, height = 0.5, color = "#D55E00") +
  xlim(2.8, 3.6) + 
  theme_minimal()
```
